% mainfile: ../main.tex
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
It's been a journey.
In a way, this endeavor started nine years ago, when I took up a HiWi job for a few months before the start of my Bachelor's project.
Hendrik, it is a testament to you, the group you've built, and to the research you pursue that I have kept coming back for my Master's and finally for my PhD projects despite external factors in my life pulling me away from Aachen.
Thank you for the opportunity, and for giving me the freedom, to grow, learn, and explore.
Quantum technology is a hugely interdisciplinary field of research, and your vision of the grand scheme of things as well as your appreciation of not only the experimental physics but also the theory, software, and engineering that come along with it made this group an immensely rewarding place to be.

Having been part of the group for so long, there are many people that I am grateful to have shared my time with at the institute.
Simon, who ad-hoc introduced me to the Bloch sphere when I stumbled into his office looking for a Bachelor's thesis, and whose curiosity and eagerness to learn about just about anything are inspiring.
Who could have ever imagined that the world of Glam Metal is so big and yet so small?\footnote{
    This was before they were famous.
}
Pascal, who supervised both my Bachelor's and Master's projects, and whose influence on both the group and my own personal growth cannot be overstated.
Julian, for sticking to theory amidst so many experimentalists and teaching them lessons in diplomacy.
René, with whom I shared the central office for the longest time, and who was my partner in solitude manning the office during the Covid years.
Alex, who took his place -- literally speaking -- and continues to teach everyone sour lessons in roasting.
Sebastian, who labored and toiled away in the clean room, and who I shared pains and successes with on the lone isle of optics land---it's been a joy collaborating with you!
Thomas, whose work ethic enabled him to do Sebastian's and my job simultaneously, and whose shoulders most of my work stands on.

I have appreciated discussions in the optics team, in particular with Beata Kardynał, who was always available to share insight into the field of semiconductor optics that was new to me, and Maxim, who is wise beyond his years.
Thank you also to the mechanical workshop, Frank and Helmut in particular, who always find a solution to whichever problem you come to them with.
I am grateful to all my proof readers -- Paul, Max, Alex, Sebastian, Evelyn, Maxim, and Dominik -- that have had to cope with my complex thesis repository layout and knack for convoluted sentences.

Pursuing a PhD is more than just a job, and so I am also thankful to the people who have made it possible to reach this point or whose paths have touched mine during the course of my time here.
Constanze and Ulrich, who took me in and gave me a roof over my head 13 years ago when I came to Aachen with a backpack and a duffle bag, surprised at the fact I had not found a place to live yet despite having looked for all but a few weeks before the start of the semester.
The anonymous gentleman walking his dogs in Melaten every morning who was a regular friendly face during the lockdown weeks and beyond.
\enquote{Einen erfolgreichen Tag und bleiben Sie gesund!}
Stefan Schael, for being a demanding but understanding neighbor.

Crucially, and emphatically unironically, my parents, who have placed me in the privileged position of being able to pursue my studies -- both of them!\@ -- without financial worries or lofty expectations looming over my head.
Papa, without your enduring and patient help and support I would have never made it this far.

It is perhaps fitting that as this chapter of my life closes, another, one that also started nine years ago, merely changes shape.
Idske, you have been with me through all the highs and lows, you have helped me make decisions, lifted my spirits, shared my joy, made me laugh---in short, you have made my life the life it is, and I will not imagine it any other way anymore.
I cannot wait to start the next part of that chapter with you.

\begin{flushright}
    \itshape
    Tobias Hangleiter\\
    Aachen, September 2025
\end{flushright}
